---
title: "FASTQ File Processing"
author: "Tianyi Xu"
date: "2023-10-02"
output: html_document
---

```{r setup, include=FALSE}
# install if not installed
install.packages("devtools")
library("devtools")
devtools::install_github("benjjneb/dada2", ref="v1.16") 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the required packages
library("dada2")
```

## Process read data

```{r}
# path
path = "data/"

sample.dirs = list.dirs(path, full.names = FALSE, recursive = FALSE)
```

```{r}
fnsF = vector("list", length(sample.dirs))
fnsR = vector("list", length(sample.dirs))

for (i in seq_along(sample.dirs)) {
    sample.dir = sample.dirs[i]
    fnsF[[i]] = list.files(file.path(path, sample.dir), pattern="_1.fastq", full.names=TRUE)
    fnsR[[i]] = list.files(file.path(path, sample.dir), pattern="_2.fastq", full.names=TRUE)
}

fnsF = sort(unlist(fnsF))
fnsR = sort(unlist(fnsR))

# find sample names

sample.names = vector("character", length(fnsF))

for (i in seq_along(fnsF)) {
    sample.names[i] = unlist(strsplit(unlist(strsplit(fnsF[i], "//"))[2], "/"))[1]
}
```

# inspect the quality of the reads

```{r}
plotQualityProfile(fnsF[1:2])
```

```{r}
plotQualityProfile(fnsR[1:2])
```



position 240 and 160 trim the reads.?

## Filter and trim reads

```{r}
filtFs = file.path(path, "filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs = file.path(path, "filtered", paste0(sample.names, "_R_filt.fastq.gz"))
names(filtFs) = sample.names
names(filtRs) = sample.names

out = filterAndTrim(fnsF, filtFs, fnsR, filtRs, truncLen=c(240,160),
                    maxN=0, maxEE=c(2,2), truncQ=2, rm.phix=TRUE,
                    compress=TRUE, multithread=FALSE)
```

```{r}
head(out)
```

Calculate error

```{r}
errF = learnErrors(filtFs, multithread=FALSE)
errR = learnErrors(filtRs, multithread=FALSE)
```

Curr output: > errF = learnErrors(filtFs, multithread=FALSE)
132468480 total bases in 551952 reads from 5 samples will be used for learning the error rates.
> errR = learnErrors(filtRs, multithread=FALSE)
102837120 total bases in 642732 reads from 6 samples will be used for learning the error rates.

```{r}
plotErrors(errF, nominalQ=TRUE)
```

## Sample Inference

```{r}
dadaFs <- dada(filtFs, err=errF, multithread=FALSE)
```

```{r}
# save runtimes
save(dadaFs, file="dadaFs.rda")
save(errF, file="errF.rda")
save(errR, file="errR.rda")
save(out, file="out.rda")
```
